\textbf{3D Finite Element Model of the Chinchilla Inner Ear for Analysis of Multisystemic Pathologies}

### 2021-06-28 paper.tex
Version 4

\begin{abstract}
The mammalian inner ear consists of the cochlea and vestibular systems. These systems act as biomechanical sensors responsible for converting acoustic waves, linear translation, and angular motion experienced by the head into sensible impulses encoded in 3D space. This paper presents current progress towards a 3D FE model of the chinchilla inner ear capable of both mechanical and electrical analysis. This includes a functional mechanical model of the inner ear with displacement evaluable at the basilar membrane, cupulas, and vestibular maculae. Geometric results of modeling the cochlear, vestibular, and facial nerves are also presented. Accurate modeling of this region is accomplished through the use of $\mu$MRI and $\mu$CT scanning technology with details verified by literature research. This follows a protocol for rapid modeling which is presented here in its current form. Results from harmonic acoustic simulation and simulation of both linear and angular acceleration are shown and compared with those received from clinical studies to verify the models integrity. \newline \newline \textbf{Keywords:} inner ear, vestibular, cochlea, chinchilla, labyrinth, finite element, mechanics
\end{abstract}

@@

The point clouds obtained from $\mu$MRI imaging were imported as .stl files into MeshMixer for smoothing of the outer surfaces as well as construction of the membranous labyrinth based on extrapolation of the imported data. Artifacts which resulted in additional unattached structures were removed leaving only the desired anatomy in place. Holes and gaps in the semicircular canals and cochlea were repaired manually to ensure proper connectivity between surfaces. Curvatures of repaired sections were matched to that of surrounding surfaces using tools native to the MeshMixer program.

\subsection{Membranous Labyrinth}
The creation of the membranous labyrinth began by creating a copy of the bony labyrinth, shrunk by a fraction to create two volumes, one enclosed within the other. The precise structure of the utricle and saccule is not concretely known due to their small size and thin membranous structure, creating high potential for damage in histological studies and difficulty in imaging for even $\mu$MRI technologies. The utricle was modeled as a slightly shrunken version of the portion of the surrounding bony labyrinth anterior to the cupulas. Cupulas structures are based off of the work conducted by Dr. Rabbitt in “Semicircular canal biomechanics in health and disease” published in 2019. \cite{rabbitt} They are modeled as spanning the majority of the cupula of each SCC.

The membranous labyrinth of the vestibular system was not effectively isolated through segmentation alone. The cross-sectional area and position of the membranous labyrinth within the SCC’s was corrected using the values stated in Iversen, M. M., Rabbitt, R. D. (2017). \cite{iversen} This was accurately accomplished by referencing individual cross section in SolidWorks, placing them appropriately within the model of the bony labyrinth, and creating a thin layer element which dynamically connects each cross section.

This accuracy will allow studies to be conducted which focus on the mechotransduction of fluid to hair cells in the ampullas. The saccule was modeled as a simple globular vesicle with ducts connected to the utricle, scala media of the cochlear membranous labyrinth, and the endolymphatic duct which is in turn connected to the endolymphatic sac. This structure will allow accurate modeling of certain diseases resulting in altered levels of endolymph production as well as capture the effect of the endolymphatic duct on wave propagation between the vestibular system and cochlea.  The diameters and paths of the ducts connecting these four vesicles were modeled based off Lo, W. W., et al.’s study titled “The Endolymphatic Duct and Sac” published in 1997. The shapes of the utricle and saccule mirror that seen in Fig. 3 of their paper, with much of their information obtained through precise histology and analysis of human cadavers, paying special attention to small cavities in petrous bone. The endolymphatic sac and duct are also included, the sac simplified as a singular vesicle of appropriate size near the boundary of the skull.\\
\vspace{0.5pt}

The cochlea was modeled primarily based on information obtained directly from MRI imaging using the previously described segmentation process. Characteristic ridges on the surface of the bony labyrinth allowed curves to be drawn along the outside of the cochleas spiraling structure. These curves were connected by planes forming a wedge whose upper component represents the boundary between the scala vestibuli and the scala media, or Reissner’s Membrane, and whose lower face represents the basilar membrane where the Organ of Corti and majority of sensory hair cells are anchored. This shape was compared with that seen in literature and was confirmed to the have correct structure. \cite{trevino:chinchmodel} \cite{ekdale:chinchanat} \cite{ni:cochmcx}

The previously described processes left many small intersections between the bony and membranous labyrinths. This problem was resolved using MeshMixer and joining the two models into a singular complex, thus allowing vertices of equal position to be created at the intersections of both structures and allowing accurate boundary conditions to be created in FE analysis.

\subsection{Nervous Tissue}
The rough geometries of the vestibular, cochlear, facial nerves, and proximal brain tissue were inserted into the completed model of the bony labyrinth and smoothed to have consistent surfaces. The attachment point between brain and nerve was left largely unaltered from original data. The model of the temporal bone, received from $\mu$CT imaging, was inserted into the model to further inform nerve and cochlear geometry. The surface of the temporal bone was smoothed and turned transparent for alteration of nervous tissue. This allowed the path and diameter of the nerve bundles to be refined, filling the tubular holes in the temporal bone captured in initial slicing. The cochlear nerve was refined by enlarging it slightly to completely fill the space within the spiral-shape of the cochlear membranous labyrinth, using Boolean subtraction of the membranous labyrinth to create a mostly uniform boundary. These efforts were compared with the work of Romano, N., Federici, M., Castaldi, A. in 2019 to ensure accuracy in nerve connectivity and path. \cite{romano:nerves}

@@

\subsection{Meshing}

\subsection{Boundary Conditions}
The outer boundaries of the mechanical model lie at the division between the perilymphatic fluid of the inner ear and the bony labyrinth. This was modeled as fixed due to the relatively small amount of information available on chinchilla temporal bone material properties. Two types of fluid–structure interactions were defined in the acoustic elements of the cochlea and vestibular system. Each surface between the lymphatic fluid and a movable structure, such as the BM, RM and RWM, was defined in the model as a fluid-structure interface (FSI) where the acoustic pressure was coupled into structural analysis. (copied from Junfeng, will alter for final paper)

\subsection{Finite Element Analysis}
-This whole section will need to be reworked
\subsubsection{Analysis of Vestibulocochlear Function}
Acoustic structural finite elemens analysis was carried out on the isolated membranous and bony labyrinths. For the purpose of this test the endolymphatic gland and sac were removed to allow better comparison to prior data for calibration of the model. The effect of these structures will be analyzed in detail in future studies. Harmonic analysis in the frequency range of X to Y Hz was conducted in Ansys Workbench 19.2 \cite{ansys}

\section{Results}
\subsection{Geometry}
A model was constructed with accurate geometries for the bony and membranous labyrinths of the chinchilla inner ear, augmented by models of the vestibulocochlear nervous system, brain tissue, and temporal bone structure. Notable additional features include the presence of the facial nerve, endolymphatic duct and sac, utricle, and saccule. Figure \ref{bonenerve} shows the course the facial nerve takes through the temporal bone, identified using CT scans which confirmed the pathways found by segmenting MRI images with corresponding lumens in the temporal bone.

Figure \ref{nvs} shows a more detailed view of the nervous system and its attachments to the membranous labyrinth. Paddle like connections can be seen on the utricle and saccule while prong connections attach to the ampulla for each semicircular canal. The cochlear nerve follows the inner shape of the cochlea and attaches throughout to the organ of corti at the basiliar membrane of the scala media. The vestibular and cochlear nerves come together into the vestibulocochlear nerve and attach to the nearby brain.
This model was tested using FE analysis to validate geometry and material properties for both the cochlea and vestibular system.

Figure \ref{wm} denotes the locations of the round and oval window membranes. By assigning different material properties and conditions to these regions proper acoustic and vestibular testing can be conducted, as was done in preliminary testing.

@@

\section{Discussion}\label{discussion}
ANALYSIS OF FE TESTING - how accurate was the model based on data, what properties/geometries could be further improved, segway into applications of the model

The accuracy of geometry in nervous tissue will allow for studies on the effect of procedures or implants on nerve function. One example is that of the installation of a cochlear implant, where electrodes could be optimized through reverse analysis of the finite element model and allow for new technologies which minimize the effect on face or tongue feeling.

Inclusion of the temporal bone will allow for studies on the effects of concussive blasts and provide an opportunity for analysis of the inner ear in its damaged state. The inclusion of the saccule, endolymphatic duct, and endolymphatic sac will allow for analysis of the effects of the cochlea on the vestibular system and vice versa. The direct relationship between hearing and balance has been found to be significant JUNFENG PAPER and is one that merits further study.

@@

Some features that are important for vestibulocochlear function have been omitted and their inclusion could improve results in the future. These include the finer details of the organ of corti, and the ductus cochlearis as well as the tectorial membrane and reticular lamina. These features would allow for more accurate analysis of the mechanics of the basiliar membrane. Another valuable addition would be a feedback system involving a seperate finite element model of a sensory hair cell. This would allow for analysis of the micromechanics of the inner ear and increase the accuracy of data of future studies into hearing and balance. \cite{choi}

@@

   The greatest effect of each system on the other can be predictably seen at the base of the cochlea and the utricular and saccular vestibules as they are the closest in proximity to one another.  Studies with rodents have confirmed this effect \cite{mizrachi}

   Testing on humans is not feasible due to the necessity for certain procedures, including electrical stimulation, which would be unethical to perform on human test subjects. Another alternative which would perhaps allow more easily transferable data would be primates however, due to limited resources coupled with this projects early stage in research, it is difficult and unethical to subject them to invasive procedures which are often easier to carry out on rodents.

This study features a finite element model constructed by coregistering $\mu$CT and $\mu$MRI scans of the chinchilla bulla. The chinchilla was chosen as the basis for this FE model as it is widely accepted as one of the closest analogs to the human inner ear due to its similar number of turns in the cochlea, structure of semicircular canals, and singular primary crista. \cite{trevino:chinchmodel} The chinchilla is perhaps a better candidate than other rodents, especially microtype varieties, as evidenced in figure \ref{response2} which shows the cochlear response of a variety of rodents and is easily compared to figure \ref{response} displaying human and chinchilla cochlear response. Only the gerbil and chinchilla exhibit the desired response and range from and in frequency. \cite{mason}
- shorten, maybe combine this with limitations section and rename

>>>>

\begin{abstract}
The mammalian inner ear consists of the cochlea and vestibular systems. These systems act as biomechanical sensors responsible for converting acoustic waves, linear translation, and angular motion experienced by the head into sensible impulses encoded in 3D space. This paper presents current progress towards a 3D FE model of the chinchilla inner ear capable of both mechanical and electrical analysis. This includes a functional mechanical model of the inner ear with response evaluable at the basilar membrane, cupulas, and vestibular maculae. Geometric results of modeling the cochlear, vestibular, and facial nerves are also presented. Accurate modeling of this region is accomplished through the use of $\mu$MRI and $\mu$CT scanning technology with details verified by literature research. This follows a protocol for rapid modeling which is presented here in its current form. Results from harmonic acoustic simulation and simulation of both linear and angular acceleration are shown and compared with those received from clinical studies to verify the models integrity. \newline \newline \textbf{Keywords:} inner ear, vestibular, cochlea, chinchilla, labyrinth, finite element, mechanics
\end{abstract}

@@

\subsection{Geometry}
The point clouds obtained from $\mu$MRI imaging were imported as .stl files into MeshMixer for smoothing of the outer surfaces as well as construction of the membranous labyrinth based on extrapolation of the imported data. Artifacts which resulted in additional unattached structures were removed leaving only the desired anatomy in place. Holes and gaps in the semicircular canals and cochlea were repaired manually to ensure proper connectivity between surfaces. Curvatures of repaired sections were matched to that of surrounding surfaces using tools native to the MeshMixer program.

The membranous labyrinth was refined in MeshMixer by creating a copy of the bony labyrinth, shrunk by a fraction to create two volumes, one enclosed within the other. The precise structure of the utricle and saccule are not concretely known due to their small size and thin membranous structure, creating high potential for damage in histological studies and difficulty in imaging for even $\mu$MRI technologies. The utricle was modeled as a slightly shrunken version of the portion of the surrounding bony labyrinth anterior to the cupulas. Cupulas structures are based off of the work conducted by Dr. Rabbitt in “Semicircular canal biomechanics in health and disease” published in 2019. \cite{rabbitt} They are modeled as spanning the majority of the cupula of each SCC. The cross-sectional area and position of the membranous labyrinth within the SCC’s was corrected using the figures shown in Iversen, M. M., Rabbitt, R. D. (2017). \cite{iversen} This was accomplished by sketching cross sections in SolidWorks, placing them appropriately within the model of the bony labyrinth smoothed in MeshMixer, and creating a thin layer element which dynamically connects each cross section. The utricle and saccule were modeled to mirror that seen in Fig. 3 of W. W., et al.’s study titled “The Endolymphatic Duct and Sac” published in 1997.

The cochlea was modeled primarily based on information obtained directly from MRI imaging using the previously described segmentation process. Characteristic ridges on the surface of the bony labyrinth allowed curves to be drawn along the outside of the cochleas spiraling structure. These curves were connected by planes forming a wedge whose superior face represents Reissner’s Membrane, and whose inferior face represents the basilar membrane. This shape was compared with that seen in literature and was confirmed to the have correct structure. \cite{trevino:chinchmodel} \cite{ekdale:chinchanat} \cite{ni:cochmcx}

talk about LENGTH CORRECTION

The rough geometries of the vestibular, cochlear, facial nerves, and proximal brain tissue were inserted into the initial model of the bony labyrinth and smoothed to have consistent surfaces. The attachment point between brain and nerve was left largely unaltered from original data. The model of the temporal bone, received from $\mu$CT imaging, was inserted into the model to further inform nerve and cochlear geometry. The cochlear nerve was refined by enlarging it to completely fill the space within the spiral-shape of the cochlear membranous labyrinth, using Boolean subtraction of the membranous labyrinth to create a mostly uniform boundary. These efforts were compared with the work of Romano, N., Federici, M., Castaldi, A. in 2019 to ensure accuracy in nerve connectivity and path. \cite{romano:nerves}

\subsection{Meshing}
Only the components relevant to the mechanical model were meshed for finite element analysis in this study. This includes the oval window membrane, round window membrane, cupulas, vestibular maculae, basilar membrane, reissner's membrane, utricle, saccule, semicircular canals, and cochlea. The mechanical model was meshed with a total of XX tetrahedral elements and xx nodes using the software Hypermesh. Solid structures used the Solid185 element type while fluids, endolymph and perilymph, were modeled using the Fluid40 element type. A higher number of elements were selected for the basilar membrane in comparison to the rest of the model, XX in total. All components were initially meshed with quadrilateral elements in 2D then split to yield a model with surface represented entirely by triangular elements. Components were assigned their respective thicknesses as seen in TABLE XXX and given proper connectivity using the imprint, extend, and create element tools native to Hypermesh.

-INCLUDE DIMENSIONS TABLE like in gans middle ear paper

@@

\subsection{Boundary Conditions}
The outer boundaries of the mechanical model lie at the division between the perilymphatic fluid of the inner ear and the bony labyrinth. This was modeled as fixed due to the relatively small amount of information available on chinchilla temporal bone material properties. Each surface between lymphatic fluid and a movable structure, such as the BM, RM and RWM, was defined in the model as a fluid-structure interface (FSI) where the acoustic pressure was coupled into structural analysis. For harmonic acoustic simulation the oval window membrane was subjected to XXXX (reference Junfeng's model to figure out what to do for analysis here). The vestibular response of the model was tested with X angular acceleration with displacement tracked for each cupula and x linear acceleration with displacement tracked for the vestibular maculae.

\section{Results}
\subsection{Geometry}
A working mechanical model model was constructed with accurate geometries for the bony and membranous labyrinths of the chinchilla inner ear with response evaluable at the basilar membrane, cupulas, and vestibular maculae. This is augmented by models of the vestibulocochlear nerve, auditory cortex, and temporal bone structure for use in later electrical analysis. Features of the mechanical model which differentiate it from others include the presence of the utricle, saccule, vestibular maculae, and geometrically accurate cochlea. A FEW FIGURES OF COMPLETED MECHANICAL MODEL, MAIN FEATURES. Figure \ref{wm} denotes the locations of the round and oval window membranes.

Figure \ref{bonenerve} shows the course the facial nerve takes through the temporal bone, identified using CT scans which confirmed the pathways found by segmenting MRI images with corresponding lumens in the temporal bone. Figure \ref{nvs} shows a more detailed view of the nervous system and its attachments to the membranous labyrinth.

@@

\section{Discussion}\label{discussion}
This study features a finite element model constructed by coregistering $\mu$CT and $\mu$MRI scans of the chinchilla bulla. The chinchilla was chosen as the basis for this FE model as it is widely accepted as one of the closest analogs to the human inner ear due to its similar number of turns in the cochlea, structure of semicircular canals, and singular primary crista. \cite{trevino:chinchmodel} The chinchilla is perhaps a better candidate than other rodents, especially microtype varieties, as evidenced in figure \ref{response2} which shows the cochlear response of a variety of rodents and is easily compared to figure \ref{response} displaying human and chinchilla cochlear response. Only the gerbil and chinchilla exhibit the desired response and range from and in frequency. \cite{mason}

   The greatest effect of the vestibular system and cochlea on the other can be predictably seen at the base of the cochlea and the utricular and saccular vestibules as they are the closest in proximity to one another.  Studies with rodents have confirmed this effect \cite{mizrachi}

@@

@@

Completed nerve, temporal bone, and auditory cortex geometries can be substituted into the model as needed later to allow for analysis of multisystemic pathologies which are the result of both mechanical and electrical malfunctions in the ear. The accuracy of geometry in nervous tissue will allow for studies on the effect of procedures or implants on nerve function. One example is that of the installation of a cochlear implant, where electrodes could be optimized through reverse analysis of the finite element model and allow for new technologies which minimize the effect on face or tongue feeling.

Some features that are important for vestibulocochlear function have been omitted and their inclusion could improve results in the future. These include the finer details of the organ of corti, the ductus cochlearis, the tectorial membrane and reticular lamina. These features would allow for more accurate analysis of the mechanics of the basilar membrane. Another valuable addition would be the involvment of a seperate finite element model representing a sensory hair cell. This would make transient analysis of multisystemic pathologies feasible by translating deflection from the mechanical model into electrical impulses which could be studied in the electrical model. \cite{choi} Inclusion of the temporal bone will allow for studies on the effects of concussive blasts and provide an opportunity for analysis of the inner ear in its damaged state. The direct relationship between hearing and balance has been found to be significant JUNFENG PAPER and is one that merits further study.
